---
output: html_document
---
# CodeBook

### Author: Salvatore Lenza 
##### (https://github.com/sas08051968/Getting-and-Cleaning-Data)
##### Saturday, July 25, 2015




## Data Set Information:

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.

## Dataset structure

```{r, echo=FALSE}
setwd("~/coursera/g_c_data_project")

if (!require("data.table")) {
  install.packages("data.table")
}
if (!require("knitr")) {
  install.packages("knitr")
}
if (!require("reshape2")) {
  install.packages("reshape2")
}

require("data.table")
require("knitr")
require("reshape2")

zip_File <- "getdata-projectfiles-UCI HAR Dataset.zip"

if(!file.exists(zip_File)) {
  zip_url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
  download.file(url=zip_url,destfile=zip_File)
}

if(!(file.exists("./UCI HAR Dataset/activity_labels.txt") && file.exists("./UCI HAR Dataset/features.txt"))) { 
  unzip(zip_File) 
}

# Load activity labels
activity_labels <- read.table("./UCI HAR Dataset/activity_labels.txt")[,2]

# Load data column names
features <- read.table("./UCI HAR Dataset/features.txt")[,2]

# Extract only the measurements on the mean and standard deviation for each measurement.
extract_features <- grepl("mean|std", features)

# Load and process X_test & y_test data.
X_test <- read.table("./UCI HAR Dataset/test/X_test.txt")
y_test <- read.table("./UCI HAR Dataset/test/y_test.txt")
subject_test <- read.table("./UCI HAR Dataset/test/subject_test.txt")

names(X_test) = features

# Extract only the measurements on the mean and standard deviation for each measurement.
X_test = X_test[,extract_features]

# Load activity labels
y_test[,2] = activity_labels[y_test[,1]]
names(y_test) = c("Activity_ID", "Activity_Label")
names(subject_test) = "subject"

# Bind data
test_data <- cbind(as.data.table(subject_test), y_test, X_test)

# Load and process X_train & y_train data.
X_train <- read.table("./UCI HAR Dataset/train/X_train.txt")
y_train <- read.table("./UCI HAR Dataset/train/y_train.txt")

subject_train <- read.table("./UCI HAR Dataset/train/subject_train.txt")

names(X_train) = features

# Extract only the measurements on the mean and standard deviation for each measurement.
X_train = X_train[,extract_features]

# Load activity data
y_train[,2] = activity_labels[y_train[,1]]
names(y_train) = c("Activity_ID", "Activity_Label")
names(subject_train) = "subject"

# Bind data
train_data <- cbind(as.data.table(subject_train), y_train, X_train)

# Merge data
data = rbind(test_data, train_data)

id_labels   = c("subject", "Activity_ID", "Activity_Label")
data_labels = setdiff(colnames(data), id_labels)
melt_data      = melt(data, id = id_labels, measure.vars = data_labels)

# Apply mean function to dataset using dcast function
tidy_data   = dcast(melt_data, subject + Activity_Label ~ variable, mean)

```

```{r}
str(tidy_data)
```

## Attribute Information:

For each record in the dataset it is provided: 
* Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration. 
*Triaxial Angular velocity from the gyroscope. 
* A 561-feature vector with time and frequency domain variables. 
* Its activity label. 
* An identifier of the subject who carried out the experiment.


## head of the dataset

```{r}
head(tidy_data, n=2L)
```


## Summary of variables

```{r}
summary(tidy_data)
```